---
title: "Data Visualization"
format:
  html:
    code-fold: true
    theme: custom.scss
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
      df_print: paged
---

## Data Visualization for the SPY

```{r,echo=FALSE,message=FALSE, warning=FALSE}
library(flipbookr)
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
```

The SPDR S&P 500 ETF (SPY) is an **exchange-traded** fund that aims to track the performance of the S&P 500 Index, which is a benchmark for the broader U.S. stock market. The fund invests in the same stocks and in the same proportion as the index, providing investors with exposure to large-cap U.S. companies across various sectors. The ETF offers a cost-effective and convenient way for investors to gain diversified exposure to the U.S. equity market. It is a popular choice for both long-term investors and traders who seek to take advantage of short-term market movements. With its high liquidity and tight bid-ask spreads, SPY is one of the most actively traded ETFs in the world, making it an attractive choice for many investors.

### The line plot visual of SPY

```{r}
spy <- getSymbols("SPY",auto.assign = FALSE, from = "2017-10-01",src="yahoo") 
spy=data.frame(spy)
spy <- data.frame(spy,rownames(spy))

colnames(spy)[7] = "date"

spy$date<-as.Date(spy$date,"%Y-%m-%d")

fig <- plot_ly(spy, x = ~date, y = ~SPY.Adjusted, type = 'scatter', mode = 'lines')

fig <- fig %>% layout(title = "SPY price since 2017")
fig

```

Since 2017, the stock price of SPY has generally been on an upward trend with some periodic volatility. The ETF reached its highest point in early 2020 before experiencing a sharp decline due to the COVID-19 pandemic. However, the ETF has since recovered and reached new all-time highs in early 2021.

Several events have affected the trend of SPY's stock price. In 2017, the stock market benefited from the enactment of tax cuts and deregulation under the Trump administration. In 2018, the market experienced a significant correction due to rising interest rates, trade tensions with China, and concerns about a potential economic slowdown. In 2020, the COVID-19 pandemic caused the market to experience one of the sharpest declines in history, with SPY dropping over 30% in just a few weeks. However, the market recovered as governments and central banks provided unprecedented monetary and fiscal stimulus.

Looking ahead, the high inflation resulting from the government relief bill force the Federal Reserve to combat inflation by significantly raising interest rates and shrink the balance sheet, thereby limiting the rising Consumer Price Index (CPI) and preventing the economy from falling into another recession. However, the interest rate hike, which has risen from nearly 0% to 4.75% currently, drove the price of SPY down by about 30% from it's peak (480 to 354), causing the ETF to officially enter a bear market in 2022. It is important to note that the effects of inflation and interest rate hikes on the stock market are complex, and other factors such as economic growth and global events may also have an impact. As such, it is essential for investors to diversify their portfolios and stay up to date on market conditions to navigate any potential market volatility.

### Candle Stick chart

```{r}
df <- data.frame(spy)
# convert column1 to a time series object
ts_adjusted <- as.ts(spy[,"SPY.Adjusted"])
ma30 = ma(ts_adjusted, 30)
ma90 = ma(ts_adjusted, 90)
ma200 = ma(ts_adjusted, 200)

df <- cbind(df, ma30, ma90, ma200)

figc <- df %>% plot_ly(x = ~date, type="candlestick",
          open = ~SPY.Open, close = ~SPY.Close,
          high = ~SPY.High, low = ~SPY.Low)%>%   
        layout(title = "SPY Stock Price Candlestick Chart",
         xaxis = list(title = "Dates"),
         yaxis = list(title = "Stock Price (USD)"))%>% 
        add_trace(x = ~date, y = ~ma30, type = "scatter", mode = "lines", name = "30-day moving average", inherit = F) %>%  
        add_trace(x = ~date, y = ~ma90, type = "scatter", mode = "lines", name = "90-day moving average", inherit = F) %>%
        add_trace(x = ~date, y = ~ma200, type = "scatter", mode = "lines", name = "200-day moving average", inherit = F)

figc
```

The candlestick chart of SPY since October 2017 provides a visual representation of the daily movement of the ETF's stock price. It allows investors to see how the moving average changed over time, highlighting periods of bullish or bearish sentiment in the market. In addition, it also reveals how significant economic events can affect the S&P 500, resulting in sudden spikes or drops in the overall stock price. By analyzing the chart, investors can gain insights into the market trend and identify potential trading opportunities

## The Stock Price Trend for the 3 major tickers

```{r}

options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)

tickers = c("TSLA","AAPL","MSFT" )
for (i in tickers){
  getSymbols(i,
             from = "2017-10-01",
             to = "2022-12-01")}

x <- list(
  title = "date"
)
y <- list(
  title = "value"
)

stock <- data.frame(TSLA$TSLA.Adjusted,
                    AAPL$AAPL.Adjusted,
                    MSFT$MSFT.Adjusted)


stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

stock$date<-as.Date(stock$Dates,"%Y-%m-%d")

```

```{r}
 
g4<- ggplot(stock, aes(x=date)) +
  geom_line(aes(y=TSLA, colour="TSLA"))+
  geom_line(aes(y=MSFT, colour="MSFT"))+
  geom_line(aes(y=AAPL, colour="AAPL"))+
   labs(
    title = "Stock Prices for the Tech Companies",
    subtitle = "From 2017-2022",
    x = "Date",
    y = "Adjusted Closing Prices")+
    guides(colour=guide_legend(title="Tech Companies")) 


ggplotly(g4) %>%
  layout(hovermode = "x")
```

The stocks of Tesla (TSLA), Microsoft (MSFT), and Apple (AAPL) are all significant when analyzing the trend of SPY. As three of the largest publicly traded companies in the world, they are all major components of the S&P 500 index, which the SPY ETF tracks. As such, their individual stock prices can have a significant impact on the performance of the index as a whole.

Analyzing the movement of TSLA, MSFT, and AAPL since October 2017 can provide valuable insights into market trends and investor sentiment. For example, the stock prices of TSLA, MSFT, and AAPL can be impacted by factors such as company financials, new product releases, and changes in the broader macroeconomic environment. In turn, changes in these three stocks can have a ripple effect throughout the S&P 500 and the broader market, influencing the performance of the SPY ETF.

Using one of the most well-known and high rewarded stock during the pandemic, TSLA, as an example, the company stock price saw a steady increace from 2017 to 2020. However, after the pandemic and influcenced by the overall hype market, the stock price of TSLA sky rocketed in 2021, increased nearly 900% from the lowest point, driven by strong financial results, increased production, and growing demand for electric vehicles. The stock's inclusion in the S&P 500 index in December 2020 also had a significant impact on its price, as it led to increased demand from index fund investors.